---
title: "01 Functions"
author: "Richard White"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_download: TRUE
---

## Base R

Functions are a recipe/template that can be reused multiple times. Base R already comes with a number of pre-packaged functions. The most simple functions are the mathematical functions.

`+` is a function that takes two values (`arguments`) and adds them together:
```{r}
1+2
```

`-` is a function that takes two arguments and subtracts them:
```{r}
3-2
```

`*` is a function that takes two arguments and multiplies them:
```{r}
4*4
```

`/` is a function that takes two arguments and divides them:
```{r}
8/2
```

These four functions are special because of how they look. i.e. `argument function argument` (e.g. `3 + 4`). In most situations, functions will look like `function()` or `function(argument1)` or `function(argument1, argument2)` or `function(argument1, argument2, argument3)` and so on, depending on how many arguments your function will take.

## Simple Addition

As a simple example, we will now write our own function that takes two numbers as arguments and adds them together. Basically, we are writing a function that performs the same as the `+` function.

```{r}
Addition <- function(argument1, argument2){
  retVal <- argument1 + argument2
  
  return(retVal)
}

Addition(3,5)
```

You can see that `Addition(3,5)` is in the format `function(argument1, argument2)`, which is the standard function format that you will be creating in R.

Now that we have created our `Addition` function, we can reuse this function for multiple arguments.

```{r}
Addition(2,3)
```

```{r}
Addition(21,3323)
```

```{r}
Addition(2231,331223)
```

We can also create a function that adds three numbers together:

```{r}
Addition3 <- function(argument1, argument2, argument3){
  retVal <- argument1 + argument2 + argument3
  
  return(retVal)
}

Addition3(3,5,7)
```

### Example

Try to create a function that adds four numbers together

```{r}
# YOUR FUNCTION GOES HERE
```

## Adding prespecified values

We can make the internal mechanisms of the functions more complicated. For example, we could have a function that takes one argument, adds 4 to it, and then returns that value:

```{r}
Add4ToValue <- function(argument){
  retVal <- argument + 4
  
  return(retVal)
}

Add4ToValue(3)
```

We can also reuse the function `Addition` that we created earlier:

```{r}
Add4ToValue <- function(argument){
  retVal <- Addition(argument, 4)
  
  return(retVal)
}

Add4ToValue(3)
```

### Example

Try to create a function that:

- Takes a number as an argument
- Doubles that number (i.e. multiplies it by two)
- Returns the doubled value

```{r}
# YOUR FUNCTION GOES HERE
```

## Asking questions

Your functions can also be used to ask questions. The most commonly used questioning function in base R is `is.na()`. This function asks if the argument is equal to `NA`.

```{r}
is.na(NA)
```

```{r}
is.na(3)
```

```{r}
is.na("hello")
```

We can make a similar function that asks if the argument is positive (i.e. greater than zero):

```{r}
IsPositive <- function(argument){
  if(argument>0){
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

```{r}
IsPositive(2)
```

```{r}
IsPositive(-2)
```

```{r}
IsPositive(0)
```

### Example

Try to create a function that:

- Takes a number as an argument
- Returns `TRUE` if the argument is greater than or equal to 10
- Returns `FALSE` otherwise

```{r}
# YOUR FUNCTION GOES HERE
```

### Example

Try to create a function that:

- Takes a vector (e.g. `c(1,2,3,4)`) as an argument
- Returns `TRUE` if the vector is longer than 5 elements
- Returns `FALSE` otherwise

Hints:

- Use the function `length()`

```{r}
# YOUR FUNCTION GOES HERE
```

## Selecting Parts Of Your Data

Functions can also be used to return parts of your data. For example, we could create a function that only returns the positive values from a vector:

```{r}
PositiveValues <- function(argument){
  retVal <- argument[argument>0]
  
  return(retVal)
}

PositiveValues(argument=c(-2,3,-4,6))
```

These returned values can be assigned to new variables and used further in your code:

```{r}
myReturnedValues <- PositiveValues(argument=c(-2,3,-4,6))

print(myReturnedValues*10)
```


### Example

Try to create a function that:

- Takes a vector (e.g. `c(,1,2,3,4)`) as an argument
- Returns the number of positive values

Hints:

- Use the function `length()`

```{r}
# YOUR FUNCTION GOES HERE
```





