---
title: "R Notebook"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

In stata, we define variables like:

```{}
global MAX_DATE=3
```

In R we define them like this:

```{r}
MAX_DATE <- 3
```

In STATA, the entire dataset is like an excel sheet, and you can only have one up at the same time. In R, you can have multiple "datasets" up at the same time:

```{r}
data1 <- data.frame(x=c(1,2,3),y=c(3,4,5))
data2 <- data.frame(x=c(10,20,30),y=c(30,40,50))

print(data1)
print(data2)
```

In stata, you refer to the variables of the open dataset. For example

```
gen bloodpressure=130
```

In R, you always need to refer to the dataset that you care about:

```{r}
data1$bloodpressure <- 130
print(data1)
print(data2)
```

# assigning a variable

In stata, we assign variables by doing this:

```
gen var1 = "a"
```

In R, we assign variables by doing this:

```{r}
var1 <- "a"
```

This is nice, because it makes it easier to see the difference between "=" (assigning) and "==" (a question). In R, we have "<-" (assigning) and "==" (a question).

# What kind of variables can we have?

```{r}
# character/string
var1 <- "a"
print(var1)

# numeric
var2 <- 3
print(var2)

# categorical/factors
var3 <- factor(3,levels=c(1,2,3),labels=c("low","medium","high"))
print(var3)

var3 <- factor(2,levels=c(1,2,3),labels=c("low","medium","high"))
print(var3)

# applying a categorical/factor to a dataframe
data1$overweight <- factor(NA, levels=c(1,2,3),labels=c("low","medium","high"))
print(data1)
# make the first woman overweight
data1$overweight[1] <- "low"
print(data1$overweight)
print(data1)
```


# vectors!
```{r}
var <- c(10,20,30,40,50,50,60,70)
print(var)

# i want the value in the 6th position
print(var[6])

# i want to change the value in the 2nd position
var[2] <- 1000
print(var)

# i want to change the value in the 1st position to 8
var[1] <- 8

# i want to make the 3rd value equal to the first value plus the second value
var[3] <- var[1] + var[2]

# i want to check that the 3rd value is equal to the first value plus the second value
print(var[3]==var[1] + var[2])

```

# control flow statements

In programs, we often want to change the flow of the program, depending on if certain things are true. We can use if statements to determine this.

```{r}
a = 3
if(a==6){
  print("a is 6")
} else {
  print("a is not 6")
}
```

Task:

- create a vector (called "d") containing 1, 2, 3, 4, 5
- find out the length of d (using `length`) ("how many post boxes we have")
- if the length of "d" is less than 3, print "short", otherwise print "long"

```{r}
d = c(1,2,3,4,5)
# OR
d = c(1:5)

print(length(d))

if(length(d)<3){
  print("short")
} else {
  print("long")
}

```

Task:

- create a vector (called "d") containing 1, 2, 3, 4, 5
- find out the length of d (using `length`) ("how many post boxes we have")
- if the length of "d" is less than 3, print "short", if the length of "d" is between 3 and 7 (inclusive) print "medium", otherwise print "long"
- HINT: use "else if"

```{r}
d = c(1,2,3,4,5)
# OR
d = c(1:5)

print(length(d))

if(length(d)<3){
  print("short")
} else if(length(d)<7){
  print("medium")
} else {
  print("long")
}
```

# subsetting data

Task:

- create a vector (called "d") containing -10, -9, ..., 9, 10
- create a new vector from "d" (called "pos") containing the values 1, ..., 10

```{r}
d = c(-10:10)
print(d>0)
pos = d[d>0]
print(pos)
```

Task:

- create a vector (called "d") containing -10, -9, ..., 9, 10
- create a new vector from "d" (called "neg") containing the values -10, ..., -1

```{r}
d = c(-10:10)
print(d<0)
neg = d[d<0]
print(neg)
```

Task:

- create a vector (called "A") containing -10, -9, ..., 9, 10
- create a vector (called "B") containing 100, ..., 150
- create two new vectors ("small" and "big") where "small" contains the shorter vector and "big" contains the longer vector


```{r}
A = c(-10:10)
B = c(100:150)

if(length(A) < length(B)){
  small = A
  big = B
} else {
  small = B
  big = A
}
```

Task:

- create a vector (called "A") containing 1, ..., 10
- create a vector (called "B") containing 101, ..., 110
- create a data.frame (called "d") containing A and B as variables (columns)
- create a data.frame (called "d5plus") from d, subsetted on where d$A>=5

```{r}
A = c(-10:10)
B = c(100:150)

if(length(A) < length(B)){
  small = A
  big = B
} else {
  small = B
  big = A
}



```




