---
title: "03 If statements"
author: "Richard White"
date: "2018-03-23"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

# control flow statements

In programs, we often want to change the flow of the program, depending on if certain things are true. We can use if statements to determine this.

```{r}
a = 3
if(a==6){
  print("a is 6")
} else {
  print("a is not 6")
}
```

Task:

- create a vector (called "d") containing 1, 2, 3, 4, 5
- find out the length of d (using `length`) ("how many post boxes we have")
- if the length of "d" is less than 3, print "short", otherwise print "long"

```{r}
d = c(1,2,3,4,5)
# OR
d = c(1:5)

print(length(d))

if(length(d)<3){
  print("short")
} else {
  print("long")
}

```

Task:

- create a vector (called "d") containing 1, 2, 3, 4, 5
- find out the length of d (using `length`) ("how many post boxes we have")
- if the length of "d" is less than 3, print "short", if the length of "d" is between 3 and 7 (inclusive) print "medium", otherwise print "long"
- HINT: use "else if"

```{r}
d = c(1,2,3,4,5)
# OR
d = c(1:5)

print(length(d))

if(length(d)<3){
  print("short")
} else if(length(d)<7){
  print("medium")
} else {
  print("long")
}
```

# subsetting data

Task:

- create a vector (called "d") containing -10, -9, ..., 9, 10
- create a new vector from "d" (called "pos") containing the values 1, ..., 10

```{r}
d = c(-10:10)
print(d>0)
pos = d[d>0]
print(pos)
```

Task:

- create a vector (called "d") containing -10, -9, ..., 9, 10
- create a new vector from "d" (called "neg") containing the values -10, ..., -1

```{r}
d = c(-10:10)
print(d<0)
neg = d[d<0]
print(neg)
```

Task:

- create a vector (called "A") containing -10, -9, ..., 9, 10
- create a vector (called "B") containing 101, ..., 150
- create two new vectors ("small" and "big") where "small" contains the shorter vector and "big" contains the longer vector


```{r}
A = c(-10:10)
B = c(101:150)

if(length(A) < length(B)){
  small = A
  big = B
} else {
  small = B
  big = A
}
```
