---
title: "R Notebook"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

In STATA, the entire dataset is like an excel sheet, and you can only have one up at the same time. In R, you can have multiple "datasets" up at the same time:

```{r}
data1 <- data.frame(x=c(1,2,3),y=c(3,4,5))
data2 <- data.frame(x=c(10,20,30),y=c(30,40,50))

print(data1)
print(data2)
```

In stata, you refer to the variables of the open dataset. For example

```
gen bloodpressure=130
```

In R, you always need to refer to the dataset that you care about:

```{r}
data1$bloodpressure <- 130
print(data1)
print(data2)
```

Task:

- create a vector (called "A") containing 1, ..., 10
- create a vector (called "B") containing 101, ..., 110
- create a data.frame (called "d") containing A and B as variables (columns)
- create a data.frame (called "d5plus") from d, subsetted on where d$A>=5

```{r}
A = c(1:10)
B = c(101:150)

d <- data.frame(A,B)
d5plus <- d[d$A>=5,]
print(d5plus)
```




