---
title: "04 Data frames"
author: "Richard White"
date: "2018-03-23"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

In STATA, the entire dataset is like an excel sheet, and you can only have one up at the same time. In R, you can have multiple "datasets" up at the same time:

```{r}
data1 <- data.frame(x=c(1,2,3),y=c(3,4,5))
data2 <- data.frame(x=c(10,20,30),y=c(30,40,50))

print(data1)
print(data2)
```

In stata, you refer to the variables of the open dataset. For example

```
gen bloodpressure=130
```

In R, you always need to refer to the dataset that you care about:

```{r}
data1$bloodpressure <- 130
print(data1)
print(data2)
```

Task:

- create a vector (called "A") containing 1, ..., 10
- create a vector (called "B") containing 101, ..., 110
- create a data.frame (called "d") containing A and B as variables (columns)
- create a data.frame (called "d5plus") from d, subsetted on where d$A>=5

```{r}
A = c(1:10)
B = c(101:110)

d <- data.frame(A,B)
d5plus <- d[d$A>=5,]
print(d5plus)
```

Task:

- in the data.frame `d` rename the two variables/columns to `var1` and `var2`
- hint, use the function `names()`

```{r}
names(d)
names(d) <- c("var1", "var2")
names(d)
```

Task:

- in the data.frame `d` create a new variable/column called `var3` which is equal to 1 if `var1`<5 and equal to `var2` otherwise

```{r}
rowsThatIWant <- d$var1<5
print(rowsThatIWant)

d[rowsThatIWant,]

d$var3 <- d$var2
# here we are only operating with vectors
d$var3[d$var1<5] <- 1

# here we are working within the data.frame format
d[d$var1<5,"var3"] <- 1

# here we are working within the data.frame format, making it more general
columnThatIWant <- "var3"
d[rowsThatIWant,columnThatIWant] <- 1
```

Task:

- create a data.frame `w` containing variables: `days of the week` (day) and `expected work hours` (expectedHours)

```{r}
days <- c("Mo","Tu","We","Th","Fr","Sa","Su")
expectedHours <- c(8,8,8,8,0,0,8)
# another way to create expectedHours (more efficent, using the function "rep")
expectedHours <- c(rep(8,4),0,0,8)

w <- data.frame("days"=days,"expectedHours"=expectedHours)
```

Task:

- inside `w`, create a new variable called `workedHours` and initialize it to `NA`

```{r}
w$workedHours <- NA
```

Task:

- inside `w`, set workedHours on Monday to 9

```{r}
# least generic
w$workedHours[1] <- 9

# bit more generic
dayThatIsMonday <- w$days=="Mo"
w$workedHours[dayThatIsMonday] <- 9

# data.frame style i.e. w[row,column]
w[dayThatIsMonday, "workedHours"] <- 9

# most generic data.frame style i.e. w[row,column]
columnThatIWant <- "workedHours"
w[dayThatIsMonday, columnThatIWant] <- 9

```





